---
title: "Weighted Cox Example 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, messages = FALSE)
rm(list=ls())
library(survival)
library(weightedKMplots)
library(tinytex)

# source revised functions (to be updated in weightedKMplots)
# source("~/Library/CloudStorage/OneDrive-MerckSharp&DohmeLLC/documents/GitHub/weightedKMplots/R/df_counting_improved.R")
# source("~/Library/CloudStorage/OneDrive-MerckSharp&DohmeLLC/documents/GitHub/weightedKMplots/R/km_calculations_helpers.R")
# source("~/Library/CloudStorage/OneDrive-MerckSharp&DohmeLLC/documents/GitHub/weightedKMplots/R/kmplotting_helpers.R")
# source("~/Library/CloudStorage/OneDrive-MerckSharp&DohmeLLC/documents/GitHub/weightedKMplots/R/weightedCox.R")

source("~/Documents/GitHub/weightedKMplots/R/df_counting_improved.R")
source("~/Documents/GitHub/weightedKMplots/R/km_calculations_helpers.R")
source("~/Documents/GitHub/weightedKMplots/R/kmplotting_helpers.R")
source("~/Documents/GitHub/weightedKMplots/R/weightedCox.R")

#remove.packages("future")
#remotes::install_version("future", version = "1.33.0")


```


Define at-risk processes:

$$S^{(k)}(\beta, t) = \sum_{j} Y_j(t) X_j^k \exp(X_j^\top \beta) ) $$ 
 Without ties the Cox score is 
 
$$ U(\beta) = \sum_{i=1}^n \int_0^\tau \left[ X_i - \frac{S^{(1)}(\beta, t)}{S^{(0)}(\beta, t)} \right] dN_i(t) $$

For Efron's approach define 

$$S^{(k)}_{D}(\beta, t) = \sum_{j \in D(t)} X_j^k \exp(X_j^\top \beta) )$$



 $$U(\beta) = \sum_{t} \sum_{l=1}^{d} \left[ X_{i_l} - \frac{S^{(1)}(\beta, t) - \frac{l-1}{d} S^{(1)}_D(\beta, t)}{S^{(0)}(\beta, t) - \frac{l-1}{d} S^{(0)}_D(\beta, t)} \right]$$ 
 





# Setup gbsg and rotterdam datasets
```{r, include=TRUE}
df <- gbsg
df$treat <- df$hormon
# rfstime is time to recurrence in days
# convert to months
df$time_months <- df$rfstime/30.4375 
df_gbsg <- df 

# Following Therneau example
gbsg_validate <-within(rotterdam,{
  rfstime <- ifelse(recur==1,rtime,dtime)
  #rfstime <- pmin(rtime,dtime)
  t_months<-rfstime/30.4375
  time_months <- t_months
  status <- pmax(recur,death)
  # censor if beyond 84 months ?
  #time_months <- pmin(t_months,84)
  #status <- ifelse(status==1 & t_months <= 84,1,0)
  # Second version (conservative per Therneau)
  ignore <- (recur==0 & death==1 & rtime < dtime)
  status2 <- ifelse(recur==1 | ignore, recur, death)
  rfstime2 <- ifelse(recur==1 | ignore, rtime,dtime)
  time_months2 <- rfstime2/30.4375
  grade3 <- ifelse(grade=="3",1,0)
  treat <- hormon
  id<-as.numeric(c(1:nrow(rotterdam)))  
  SG0 <- ifelse(er<=0,0,1)
})
# Check subject 40 who was censored at recurrence but death afterwards
# Event at 6.6 years
#subset(gbsg_validate,pid==40)
# node positive to match gbsg?
df <- subset(gbsg_validate, nodes > 0)
fit.ps <- glm(treat ~ age+meno+size+grade3+nodes+pgr+chemo+er, data = df, family="binomial")
pihat <- fit.ps$fitted
pihat.null <- glm(treat ~ 1, family = "binomial", data = df)$fitted
wt.1 <- pihat.null/pihat
wt.0 <- (1-pihat.null)/(1-pihat)
# Stabilized weight
df$sw.weights <- ifelse(df$treat == 1, wt.1,wt.0)

df_rotterdam <- df

tte.name<-c("time_months")
event.name<-c("status")
treat.name<-c("treat")
arms <- c("treat","control")

# Weighting for rotterdam data 
wt.name <- c("sw.weights")

```




```{r}
dfcount <- get_dfcounting(df=df_gbsg, tte.name=tte.name, event.name=event.name, treat.name=treat.name, arms=arms, by.risk=24)


atpoints <- Y
km.pool<-survfit(Surv(Y,Delta)~1)
S.pool<-summary(km.pool,c(atpoints))$surv

# Modestly weighted t*= 6, 12, 16
mb12<-wt.rg.S(S=S.pool,rho=-11,gamma=-11,tpoints=atpoints,details=FALSE,mb_tstar=12)
mb6<-wt.rg.S(S=S.pool,rho=-11,gamma=-11,tpoints=atpoints,details=FALSE,mb_tstar=6)
mb16<-wt.rg.S(S=S.pool,rho=-11,gamma=-11,tpoints=atpoints,details=FALSE,mb_tstar=16)

# FH(0,1)
wt.01<-wt.rg.S(S=S.pool,rho=0,gamma=1,tpoints=atpoints,details=FALSE)
# exp(FH(0.5,0.5))
e55 <- wt.rg.S(S=S.pool,rho=-5,gamma=-5,tpoints=atpoints,details=FALSE)

#wmax <- max(wt.new)
#wt.new2 <- pmin(exp(wt.01),wmax)

eV15 <- wt.rg.S(S=S.pool,rho=-6,gamma=-6,tpoints=atpoints,details=FALSE)

par(mfrow=c(1,1))

plot(atpoints,mb12,xlab="time",ylab="w(t)",type="s",lty=1,col="grey",lwd=3,ylim=c(0.9,2.5))
lines(atpoints,mb6,type="s",lty=2,col="grey",lwd=3)
lines(atpoints,mb16,type="s",lty=3,col="grey",lwd=3)
lines(atpoints,exp(wt.01),type="s",lty=1,col="green",lwd=3)
lines(atpoints,eV15,type="s",lty=2,col="brown",lwd=3)
lines(atpoints,e55,type="s",lty=2,col="blue",lwd=3)

legend(1,2.5,c("mb(12)","mb(6)","mb(16)","e01=exp(FH(0,1))","e55=exp(FH(0.5,0.5))","eV15=min(e01,max(e55))"),lty=c(1,2,3,1,2,2),col=c("grey","grey","grey","green","blue","brown"),lwd=3,bty="n")

  
```

```{r, fig.width = 8, fig.height=6}
plot_weighted_km(dfcount=dfcount, conf.int=TRUE, show.logrank = TRUE, ymax = 1.05)
title(main="GBSG")
```

# FH(0,0)
```{r}


#get_rg <- cox_rhogamma(dfcount = dfcount, scheme = "fh", scheme_params = list(rho = 0, gamma = 0), draws = 0, verbose = TRUE)

get_rg <- cox_rhogamma(dfcount = dfcount, scheme = "fh", scheme_params = list(rho = 0, gamma = 0), draws = 1000, verbose = TRUE)

temp <- wlr_dhat_estimates(dfcounting=dfcount, rho = 0, gamma = 0,tzero=24)
cat("Correlation between WLR and KM difference at 24-months",c(temp$cor_wlr_dhat),"\n")
temp$lr/sqrt(temp$sig2_lr)

get_rg$z.score


# logrank rho=5,gamma=0
# simtrial
# dfa <- df
# dfa$tte <- dfa$time_months
# dfa$event <- dfa$status
# dfa$treatment <- ifelse(dfa$treat ==1, "experimental", "control")
# dfa$stratum <- "All"
# dfa2 <- as.data.table(dfa) 
# check <- dfa2 |> wlr(weight = fh(rho = 1, gamma = 0))
# cat("Mine and simtrial",c(my_wlr,check$z),"\n")

# Therneau's example    
# result_fh <- cph.weighted(
#   df = df,
#   scheme = "fh",
#   scheme_params = list(rho = 0, gamma = 0),
#   outcome.name = "time_months",
#   event.name = "status",
#   treat.name = "hormon"
# )
# print(result_fh)
```

# Schemper 

```{r}
get_rg <- cox_rhogamma(dfcount = dfcount, scheme = "schemper", scheme_params = list(Scensor = dfcount$survG_all), draws = 1000, verbose = TRUE)
```

# MB
```{r}
get_rg <- cox_rhogamma(dfcount = dfcount, scheme = "MB", scheme_params = list(mb_tstar = 12), draws = 1000, verbose = TRUE)
```

# fh_exp2 
```{r}
get_rg <- cox_rhogamma(dfcount = dfcount, scheme = "fh_exp2", draws = 1000, verbose = TRUE)
```


# Simulation setup (from Keaven's examples)

```{r}

library(gsDesign)
library(gsDesign2)
library(dplyr)
library(tibble)
library(gt)
library(ggplot2)
#library(cowplot)
library(simtrial)
library(tidyr)
library(future.batchtools)
library(doFuture)
library(foreach)
library(tictoc)
```


```{r}

set.seed(8316951)

survival_at_24_months <- 0.35
hr <- log(.35)/log(.25)
control_median <- 12
control_rate <- c(log(2) / control_median, (log(.25) - log(.2)) / 12)

scenarios <- tribble(
  ~Scenario, ~Name,           ~Period, ~duration, ~Survival,
  0,         "Control",       0,       0,         1,
  0,         "Control",       1,       24,        .25,
  0,         "Control",       2,       12,        .2,
  1,         "PH",            0,       0,         1,
  1,         "PH",            1,       24,        .35,
  1,         "PH",            2,       12,        .2^hr,
  2,         "3-month delay", 0,       0,         1,
  2,         "3-month delay", 1,       3,         exp(-3 * control_rate[1]),
  2,         "3-month delay", 2,       21,        .35,
  2,         "3-month delay", 3,       12,        .2^hr,
  3,         "6-month delay", 0,       0,         1,
  3,         "6-month delay", 1,       6,         exp(-6 * control_rate[1]),
  3,         "6-month delay", 2,       18,        .35,
  3,         "6-month delay", 3,       12,        .2^hr,
  4,         "Crossing",      0,       0,         1,
  4,         "Crossing",      1,       3,         exp(-3 * control_rate[1] * 1.3),
  4,         "Crossing",      2,       21,        .35,
  4,         "Crossing",      3,       12,        .2^hr,
  5,         "Weak null",     0,       0,         1,
  5,         "Weak null",     1,       24,        .25,
  5,         "Weak null",     2,       12,        .2,
  6,         "Strong null",   0,       0,         1,         
  6,         "Strong null",   1,       3,         exp(-3 * control_rate[1] * 1.5),
  6,         "Strong null",   2,       3,         exp(-6 * control_rate[1]),
  6,         "Strong null",   3,       18,        .25,
  6,         "Strong null",   4,       12,        .2,
  )
# scenarios |> gt()

fr <- 
  scenarios |>
    group_by(Scenario) |>
  #  filter(Scenario == 2) |>
    mutate(Month = cumsum(duration),
           x_rate = -(log(Survival) - log(lag(Survival, default = 1))) / 
           duration, 
           rate = ifelse(Month > 24, control_rate[2], control_rate[1]),
           hr = x_rate / rate) |>
    select(-x_rate) |>
    filter(Period > 0, Scenario > 0) |> ungroup()
#fr |> gt() |> fmt_number(columns = everything(), decimals = 2)

fr <- fr |> mutate(fail_rate = rate, dropout_rate =0.001, stratum = "All")

# MWLR
mwlr <- fixed_design_mb(
  tau = 12,
  enroll_rate = define_enroll_rate(duration = 12, rate = 1),
  fail_rate = fr |> filter(Scenario == 2),
  alpha = 0.025, power = .85, ratio = 1,
  study_duration = 36
) |> to_integer()


er <- mwlr$enroll_rate


fail_rate <- data.frame(stratum = rep("All", 2 * nrow(fr)),
period = rep(fr$Period, 2),treatment = c(rep("control", nrow(fr)), 
rep("experimental", nrow(fr))),duration = rep(fr$duration, 2),
rate = c(fr$rate, fr$rate * fr$hr)
)

fr$stratum <- "All"
# Constant dropout rate for both treatment arms and all scenarios
dropout_rate <- data.frame(stratum = rep("All", 2), 
                           period = rep(1, 2),
                           treatment = c("control", "experimental"),
                           duration = rep(100, 2),
                           rate = rep(.001, 2)
                          )
```

# Strong null
```{r}
# strong null
dgm <- fr[c(14:17),]

fail_rate <- data.frame(stratum = rep("All", 2 * nrow(dgm)),
                          period = rep(dgm$Period, 2),
                          treatment = c(rep("control", nrow(dgm)), 
                                        rep("experimental", nrow(dgm))),
                          duration = rep(dgm$duration, 2),
                          rate = c(dgm$rate, dgm$rate * dgm$hr)
                         )
    
# Constant dropout rate for both treatment arms and all scenarios
dropout_rate <- data.frame(stratum = rep("All", 2), 
                           period = rep(1, 2),
                           treatment = c("control", "experimental"),
                           duration = rep(100, 2),
                           rate = rep(.001, 2)
                          )

er <- mwlr$enroll_rate

```



```{r}
# Generate a dataset
dat <- sim_pw_surv(n = 698, enroll_rate = er,
fail_rate = fail_rate, dropout_rate = dropout_rate)
analysis_data <- cut_data_by_date(dat, 36)
dfa <- analysis_data
dfa$treat <- ifelse(dfa$treatment=="experimental",1,0)
```

```{r}
dfcount <- get_dfcounting(df=dfa, tte.name = "tte", event.name = "event", treat.name = "treat", arms = "treatment", by.risk= 6)
```

```{r, fig.width = 8, fig.height=6}
plot_weighted_km(dfcount=dfcount, conf.int=TRUE, show.logrank = TRUE, ymax = 1.05)
title(main="Simulated data")
```


# Correlation between log-rank and milestone (24-months) (tzero=24)
```{r}

temp <- wlr_dhat_estimates(dfcounting=dfcount, rho = 1, gamma = 0,tzero=24)
cat("Correlation between WLR and KM difference at 24-months",c(temp$cor_wlr_dhat),"\n")
my_wlr <- temp$lr/sqrt(temp$sig2_lr)
# logrank rho=5,gamma=0
# simtrial
check <- dfa |> wlr(weight = fh(rho = 1, gamma = 0))
cat("Mine and simtrial",c(my_wlr,check$z),"\n")

coxz <- cox_rhogamma(dfcount = dfcount, scheme = "fh", scheme_params = list(rho = 1, gamma = 0), draws = 0, verbose = TRUE)$z.score
coxz
coxz^2

temp <- survdiff(Surv(tte,event) ~ treat, data=dfa, rho = 1)
temp$chisq
temp$pvalue

temp <- dfa |> wlr(weight = mb(delay = 6, w_max = Inf))
mb6z <- temp$z
mb6z 

```

# FH(0,1)
```{r}
get_rg <- cox_rhogamma(dfcount = dfcount, scheme = "fh", scheme_params = list(rho = 0, gamma = 1), draws = 1000, verbose = TRUE)
```

# MB
```{r}
get_rg <- cox_rhogamma(dfcount = dfcount, scheme = "MB", scheme_params = list(mb_tstar = 12), draws = 1000, verbose = TRUE)
```

# fh_exp2 
```{r}
get_rg <- cox_rhogamma(dfcount = dfcount, scheme = "fh_exp2", draws = 1000, verbose = TRUE)
```

```{r}
sim_fn <- function(scen, enroll_rate, dropout_rate, fr, delay=12){

fail_rate <- data.frame(stratum = rep("All", 2 * nrow(fr)),
period = rep(fr$Period, 2),treatment = c(rep("control", nrow(fr)), 
rep("experimental", nrow(fr))),duration = rep(fr$duration, 2),
rate = c(fr$rate, fr$rate * fr$hr)
)
# Generate a dataset
dat <- sim_pw_surv(n = 698, enroll_rate = enroll_rate,
fail_rate = fail_rate, dropout_rate = dropout_rate)

analysis_data <- cut_data_by_date(dat, 36)
dfa <- analysis_data
dfa$treat <- ifelse(dfa$treatment=="experimental",1,0)

maxcombopz <- (analysis_data |> maxcombo(rho = c(0,0), gamma = c(0,0.5), return_corr = TRUE))
 logrankz <- maxcombopz$z[1]
 fh05z <- maxcombopz$z[2]

 maxcombopz2 <- (analysis_data |> maxcombo(rho = c(0,0), gamma = c(0,1), return_corr = TRUE))
 fh01z <- maxcombopz2$z[2]

  maxcombop <- maxcombopz$p_value
  
  mb6z <- (analysis_data |> wlr(weight = mb(delay = 6, w_max = Inf)))$z
  
  mb12z <- (analysis_data |> wlr(weight = mb(delay = 12, w_max = Inf)))$z
  
  mb16z <- (analysis_data |> wlr(weight = mb(delay = 16, w_max = Inf)))$z
  
  # Mine
  # Note: turnoff KM and seKM checking so that plots are not called within parallel sessions
  dfcount <- get_dfcounting(df=dfa, tte.name = "tte", event.name = "event", treat.name = "treat", arms = "treatment", by.risk= 6, check.KM = FALSE, check.seKM = FALSE)
  # MB
  mb12z_mine <- cox_rhogamma(dfcount = dfcount, scheme = "MB", scheme_params = list(mb_tstar = 12))$z.score
 
 fhexp2z <- cox_rhogamma(dfcount = dfcount, scheme = "fh_exp2")$z.score

 fh01z_mine <- cox_rhogamma(dfcount = dfcount, scheme = "fh", scheme_params = list(rho = 0, gamma = 1))$z.score

 
    return(data.frame(Scenario = scen, 
                    MaxCombo = maxcombop,
                    Logrank = logrankz,
                    FH01 = fh01z,
                    FH05 = fh05z,
                    MB6 = mb6z,
                    MB12 = mb12z,
                    MB16 = mb16z,
                    MB12mine = mb12z_mine,
                    FHexp2 = fhexp2z,
                    FH01mine = fh01z_mine
                    ))
}

```


```{r}

#options(future.makeNodePSOCK.setup_strategy = "sequential")

#plan(multisession(workers = 10))

#plan(multicore)
# also run as R script
#Rscript script.R

library(future.callr)
plan(callr)

set.seed(8316951)

# 50K @ 3.5 hrs

n_sim <- 100*1000

fr$stratum <- "All"
# Constant dropout rate for both treatment arms and all scenarios
dropout_rate <- data.frame(stratum = rep("All", 2), 
                           period = rep(1, 2),
                           treatment = c("control", "experimental"),
                           duration = rep(100, 2),
                           rate = rep(.001, 2)
                          )
tictoc::tic()
# Do simulations by scenario
results_sims <- foreach(scen = 1:6, .combine = 'rbind')%:%
  foreach(sim = 1:n_sim, .combine = 'rbind', .options.future = list(seed = TRUE)) %dofuture% {
    library(simtrial)
tryCatch({sim_fn(scen = scen, enroll_rate = mwlr$enroll_rate, dropout_rate = dropout_rate, 
           fr = fr |> dplyr::filter(Scenario == scen))},
         error = function(e) NA)
  }

plan(sequential)
tictoc ::toc()

save(results_sims, file="results/sims_example1_mac_100k.Rdata")

```


# Weighted log-rank statistics


```{r}

results <- res_out$results_sims
results$Scenario <- factor(results$Scenario, levels = 1:6,
     labels = c("PH", "3m delay", "6m delay", "Crossing", "Weak null", "Strong null"))
results |> group_by(Scenario) |> 
  summarise("logrank" = mean(logrankz > qnorm(0.975)),
            "fh00*" = mean(fh00z_mine > qnorm(0.975)),  
            "FH05" = mean(fh05z > qnorm(0.975)),
            "FH05*" = mean(fh05z_mine > qnorm(0.975)),
            "FH01" = mean(fh01z > qnorm(0.975)),
            "FH01*" = mean(fh01z_mine > qnorm(0.975)),
            "FH01db" = mean(fh01z_debiased > qnorm(0.975)),
            "MB(6)" = mean(mb6z > qnorm(0.975)),
            "MB(6)*" = mean(mb6z_mine > qnorm(0.975)),
            "MB(12)" = mean(mb12z > qnorm(0.975)),
            "MB(12)*" = mean(mb12z_mine > qnorm(0.975)),
            "MB(16)" = mean(mb16z > qnorm(0.975)),
            "MB(16)*" = mean(mb16z_mine > qnorm(0.975)),
            "FHexp1" = mean(fhe1z > qnorm(0.975)),
            "FHexp2" = mean(fhe2z > qnorm(0.975)),
            "maxcombbo" = mean(maxcombop <= 0.025)
            ) |>
  gt() |> fmt_number(columns = 2:16, decimals = 3)  

```


# WLR and Wald tests (upper-bound < 1)

```{r}

results <- res_out$results_sims
results$Scenario <- factor(results$Scenario, levels = 1:6,
     labels = c("PH", "3m delay", "6m delay", "Crossing", "Weak null", "Strong null"))
results |> group_by(Scenario) |> 
  summarise("Cox" = mean(fh00_wald < 1.0),
            "Cox*" = mean(fh00_waldstar < 1.0),
            "fh00z" = mean(fh00z_mine > qnorm(0.975)),  
            "FH05w" = mean(fh05_wald < 1.0),
            "FH05w*" = mean(fh05_waldstar < 1.0),
            "FH05z" = mean(fh05z_mine > qnorm(0.975)),
            "FH01w" = mean(fh01_wald < 1.0),
            "FH01w*" = mean(fh01_waldstar < 1.0),
            "FH01z" = mean(fh01z_mine > qnorm(0.975)),
            "MB(12)w" = mean(mb12_wald < 1.0),
            "MB(12)w*" = mean(mb12_waldstar < 1.0),
            "MB(12)z" = mean(mb12z_mine > qnorm(0.975)),
            "FHe1w" = mean(fhe1_wald < 1.0),
            "FHe1w*" = mean(fhe1_waldstar < 1.0),
            "FHe1z" = mean(fhe1z > qnorm(0.975)),
            "FHe2w" = mean(fhe2_wald < 1.0),
            "FHe2w*" = mean(fhe2_waldstar < 1.0),
            "FHe2z" = mean(fhe2z > qnorm(0.975))
            ) |>
  gt() |> fmt_number(columns = 2:18, decimals = 3)  

```
       
       
       
       
# Hazard-ratio estimates

```{r}

results <- res_out$results_sims
results$Scenario <- factor(results$Scenario, levels = 1:6,
     labels = c("PH", "3m delay", "6m delay", "Crossing", "Weak null", "Strong null"))
results |> group_by(Scenario) |> 
  
    summarise("Cox" = mean(exp(fh00_bhat), na.rm=TRUE),
              "Coxdb" = mean(exp(fh00_bhatdebiased), na.rm=TRUE),
              "se00" = sqrt(var(fh00_bhat, na.rm=TRUE)),
              "se00h" = mean(fh00_sigbhat, na.rm=TRUE),
              "se00h*" = mean(fh00_sigbhatstar, na.rm = TRUE),
              "FH01" = mean(exp(fh01_bhat), na.rm=TRUE),
              "FH01db" = mean(exp(fh01_bhatdebiased), na.rm=TRUE),
              "se01" = sqrt(var(fh01_bhat, na.rm=TRUE)),
              "se01h" = mean(fh01_sigbhat, na.rm=TRUE),
              "se01h*" = mean(fh01_sigbhatstar, na.rm = TRUE)
            ) |>
  gt() |> fmt_number(columns = 2:11, decimals = 3)  

```

       


## Simulations

Simulations below are based on `r prettyNum(n_sim, big.mark=",")` trials simulated for each result. Type-1 error upper bounds are `r round(0.025 + 1.96*sqrt(0.025*0.975/n_sim),4)` and 
`r round(0.05 + 1.96*sqrt(0.05*0.95/n_sim),4)` for $2.5\%$ (1-sided) and $5\%$ (2-sided) alpha-levels.  


